---
title: "DSC_520_week10_Assignment00"
author: "Reenie Christudass"
date: "`r Sys.Date()`"
output:
  pdf_document: default
  html_document: default
---

\tableofcontents
\pagebreak

## Load libraries as needed

```{r}
if(!require('factoextra')) {
  install.packages("factoextra", repos="http://cran.us.r-project.org")
  library('factoextra')
}
if(!require('cluster')) {
  install.packages("cluster", repos="http://cran.us.r-project.org")
  library('cluster')
}
if(!require('NbClust')) {
  install.packages("NbClust", repos="http://cran.us.r-project.org")
  library('NbClust')
}
if(!require('e1071')) {
  install.packages("e1071", repos="http://cran.us.r-project.org")
  library('e1071')
}
if(!require('caTools')) {
  install.packages("caTools", repos="http://cran.us.r-project.org")
  library('caTools')
}
if(!require('class')) {
  install.packages("class", repos="http://cran.us.r-project.org")
  library('class')
}
if(!require('dplyr')) {
  install.packages("dplyr", repos="http://cran.us.r-project.org")
  library('dplyr')
}
```

```{r}
rm(list = ls())
## Set the working directory to the root of your DSC 520 directory
setwd("C:/Users/chris/dsc520/data")
```

## Read the binary-classifier-data
```{r}
df_binary <- read.csv("C:/Users/chris/dsc520/data/binary-classifier-data.csv")
head(df_binary)
df_binary_copy <- df_binary
```



## Euclidean distance of binary-classifier-data
```{r}
#calculate Euclidean distance between columns a and d
euclidean <- function(x, y) sqrt(sum((x - y)^2))

print("Euclidean distance (binary-classifier-data) between two columns x and y in a dataframe is: ")

euclidean(df_binary$x, df_binary$y)

```

### Convert y to be a factor
```{r}
is.factor(df_binary$y)
df_binary$y <- as.factor(df_binary$y)
head(df_binary)
```


## Build the model for binary-classifier-data
```{r}
binary_model <- glm(y~x, data=df_binary,family="binomial")
summary (binary_model)
#abline(binary_model, col='blue')
```
## Predict using model
```{r}
df_binary$Predicted <- predict(binary_model, df_binary, type='response')
head(df_binary$Predicted )
```
### convert predicted values back to true/false
```{r}
df_binary$Predicted_T_F  <- ifelse (df_binary$Predicted >= 0.5, "true", "false")
head(df_binary)
```
## Determine accuracy of the model
```{r}
accuracy <- mean(df_binary$Predicted == df_binary$y)
print(accuracy)
## ACCURACY is 0
```

## Plot 
```{r}
# Scatterplot
plot(x = df_binary$x,
     y = df_binary$y,
     xlab = "x",
     ylab = "y",
     main = "Adding a regression line with abline()"
     )

# Calculate regression model
model <- lm(formula = y ~ x,
                  data = df_binary)

# Add regression line
abline(model,
       col = "red", lwd = 2)
```





## Read the trinary-classifier-data
```{r}
df_trinary <- read.csv("C:/Users/chris/dsc520/data/trinary-classifier-data.csv")
head(df_trinary)

```

## Euclidean distance of trinary-classifier-data
```{r}
#calculate Euclidean distance between columns a and d
euclidean <- function(x, y) sqrt(sum((x - y)^2))

print("Euclidean distance (trinary-classifier-data) between two columns x and y in a dataframe is: ")

euclidean(df_trinary$x, df_trinary$y)

```
### Convert y to be a factor
```{r}
is.factor(df_trinary$y)
df_trinary$y <- as.factor(df_trinary$y)
head(df_trinary)
```

## Build the model for binary-classifier-data
```{r}
trinary_model <- glm(y~x, data=df_trinary,family="binomial")
summary (trinary_model)
#abline(binary_model, col='blue')
```
## Predict using model
```{r}
df_trinary$Predicted <- predict(trinary_model, df_trinary, type='response')
head(df_trinary)
```
### convert predicted values back to true/false
```{r}
df_trinary$predict_T_F <- ifelse (df_trinary$Predicted >= 0.5, "true", "false")
head(df_trinary)
```
## Determine accuracy of the model
```{r}
accuracy <- mean(df_trinary$y == df_trinary$Predicted)
print(accuracy)
## ACCURACY is 0
```
## Plot 
```{r}
# Scatterplot
plot(x = df_trinary$x,
     y = df_trinary$y,
     xlab = "x",
     ylab = "y",
     main = "Adding a regression line with abline()"
     )

# Calculate regression model
model <- lm(formula = y ~ x,
                  data = df_trinary)

# Add regression line
abline(model,
       col = "red", lwd = 2)
```


